---
title: "Project 1: Environment and Society"
always_allow_html: yes
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, fig.align = "center", warning = F, message = F, R.options=list(max.print=100))
```

## Hye Jeong Lee (HL9563)

### Step 1: Explore Data/ Introduction
```{r}
# load data
library(readxl)

#import downloaded data No.1
se <- read.csv("SustainableEnergy.csv")
esg <- read.csv("ESG.csv")

#get a glimpse of each data set
head (se)
head (esg)
```
*The purpose of this project is to compare the data and see if there are strong correlation between any environmental-related variables and social-related variables, and also to see if there is any correlation with GDP. The first data set se is obtained from Sustainable Energy for All, which has data related to energy. The second data set esg is obtained from Environment Social and Governance (ESG) data, which contains socioeconomic factors and other environment-related variables.* 

*I find them interesting because there is a strong bias about how living a sustainable lifestyle costs more and it is for people with money, even though it is not true. I also find them interesting because it is hard to visualize how sustainability affects social factors positively. I hope to find some correlation between environmental-related variables and social-related variables to promote for more renewable energy, and no strong correlation between environmental-related variables and GDP.*


__From se data:__

1. access_cft_tot: access to clean fuel and technologies for cooking (% of total popluation)

2. access_electricity_tot:Access to electricity (% of total population)

3. ren_electricity_output: Renewable electricity output (GWh) 

4. share_re_in_electricity: Renewable electricity share of total electricity output (%)

5. re_consumption:Renewable energy consumption (TJ)

6. total_electricity_output:Total electricity output (GWh)

7. share_total_re_in_TEFC: Renewable energy share of TFEC (%) 

8. total_final_energy_consum: Total final energy consumption (TFEC) (TJ)

9. primary_energy_intensity:Energy intensity level of primary energy (MJ/2011 USD PPP)


__From esg data:__

1. Adjusted savings: net forest depletion (% of GNI) - Net forest depletion is calculated as the product of unit resource rents and he excess of roundwood harvest over natural growth.

2. Annual freshwater withdrawals, total (% of internal resources) - total water withdrawals, not counting evaporation losses from storage basins. Withdrawals also include water from desalination plants in countries where they are a significant source. Withdrawals can exceed 100 percent of total renewable resources where extraction from nonrenewable aquifers or desalination plants is considerable or where there is significant water reuse. Withdrawals for agriculture and industry are total withdrawals for irrigation and livestock production and for direct industrial use (including withdrawals for cooling thermoelectric plants). Withdrawals for domestic uses include drinking water, municipal use or supply, and use for public services, commercial establishments, and homes. Data are for the most recent year available for 1987-2002.

3. Agricultural land (% of land area) - the share of land area that is arable, under permanent crops, and under permanent pastures. 

4. Arable land includes land defined by the FAO as land under temporary crops (double-cropped areas are counted once), temporary meadows for mowing or for pasture, land under market or kitchen gardens, and land temporarily fallow. Land abandoned as a result of shifting cultivation is excluded. Land under permanent crops is land cultivated with crops that occupy the land for long periods and need not be replanted after each harvest, such as cocoa, coffee, and rubber. This category includes land under flowering shrubs, fruit trees, nut trees, and vines, but excludes land under trees grown for wood or timber. Permanent pasture is land used for five or more years for forage, including natural and cultivated crops.

5. Agriculture, value added (% of GDP) - includes forestry, hunting, and fishing, as well as cultivation of crops and livestock production. Value added is the net output of a sector after adding up all outputs and subtracting intermediate inputs. It is calculated without making deductions for depreciation of fabricated assets or depletion and degradation of natural resources. The origin of value added is determined by the International Standard Industrial Classification (ISIC), revision 3. Note: For VAB countries, gross value added at factor cost is used as the denominator.

6. CO2 emissions (metric tons per capita) - those stemming from the burning of fossil fuels and the manufacture of cement. They include carbon dioxide produced during consumption of solid, liquid, and gas fuels and gas flaring.

7. Cooling Degree Days (projected change in degree Celsius) - measurement designed to quantify the demand for energy needed to cool buildings. It is the number of degrees that a day's average temperature is above 18°C.

8. Control of Corruption: Estimate - perceptions of the extent to which public power is exercised for private gain, including both petty and grand forms of corruption, as well as capture" of the state by elites and private interests. Estimate gives the country's score on the aggregate indicator

9. Droughts, floods, extreme temperatures (% of population, average 1990-2009) - annual average percentage of the population that is affected by natural disasters classified as either droughts, floods, or extreme temperature events. A drought is an extended period of time characterized by a deficiency in a region's water supply that is the result of constantly below average precipitation. A drought can lead to losses to agriculture, affect inland navigation and hydropower plants, and cause a lack of drinking water and famine. A flood is a significant rise of water level in a stream, lake, reservoir or coastal region. Extreme temperature events are either cold waves or heat waves. A cold wave can be both a prolonged period of excessively cold weather and the sudden invasion of very cold air over a large area. Along with frost it can cause damage to agriculture, infrastructure, and property. A heat wave is a prolonged period of excessively hot and sometimes also humid weather relative to normal climate patterns of a certain region. Population affected is the number of people injured, left homeless or requiring immediate assistance during a period of emergency resulting from a natural disaster; it can also include displaced or evacuated people. Average percentage of population affected is calculated by dividing the sum of total affected for the period stated by the sum of the annual population figures for the period stated.

10. Electricity production from coal sources (% of total) - Sources of electricity refer to the inputs used to generate electricity. Coal refers to all coal and brown coal, both primary (including hard coal and lignite-brown coal) and derived fuels (including patent fuel, coke oven coke, gas coke, coke oven gas, and blast furnace gas). Peat is also included in this category.
Energy imports, net (% of energy use) - Net energy imports are estimated as energy use less production, both measured in oil equivalents. A negative value indicates that the country is a net exporter. Energy use refers to use of primary energy before transformation to other end-use fuels, which is equal to indigenous production plus imports and stock changes, minus exports and fuels supplied to ships and aircraft engaged in international transport.

11. Energy intensity level of primary energy (MJ/$2011 PPP GDP) - the ratio between energy supply and gross domestic product measured at purchasing power parity. Energy intensity is an indication of how much energy is used to produce one unit of economic output. Lower ratio indicates that less energy is used to produce one unit of output.

12. Energy use (kg of oil equivalent per capita) - use of primary energy before transformation to other end-use fuels, which is equal to indigenous production plus imports and stock changes, minus exports and fuels supplied to ships and aircraft engaged in international transport.

13. Fertility rate, total (births per woman) - the number of children that would be born to a woman if she were to live to the end of her childbearing years and bear children in accordance with age-specific fertility rates of the specified year.
Food production index (2004-2006 = 100) - covers food crops that are considered edible and that contain nutrients. Coffee and tea are excluded because, although edible, they have no nutritive value.

14. Forest area (% of land area) - Forest area is land under natural or planted stands of trees of at least 5 meters in situ, whether productive or not, and excludes tree stands in agricultural production systems (for example, in fruit plantations and agroforestry systems) and trees in urban parks and gardens.

15. GDP growth (annual %) - Annual percentage growth rate of GDP at market prices based on constant local currency. Aggregates are based on constant 2010 U.S. dollars. GDP is the sum of gross value added by all resident producers in the economy plus any product taxes and minus any subsidies not included in the value of the products. It is calculated without making deductions for depreciation of fabricated assets or for depletion and degradation of natural resources.

16. Fossil fuel energy consumption (% of total) - Fossil fuel comprises coal, oil, petroleum, and natural gas products.

17. GHG net emissions/removals by LUCF (Mt of CO2 equivalent) - GHG net emissions/removals by LUCF refers to changes in atmospheric levels of all greenhouse gases attributable to forest and land-use change activities, including but not limited to (1) emissions and removals of CO2 from decreases or increases in biomass stocks due to forest management, logging, fuelwood collection, etc.; (2) conversion of existing forests and natural grasslands to other land uses; (3) removal of CO2 from the abandonment of formerly managed lands (e.g. croplands and pastures); and (4) emissions and removals of CO2 in soil associated with land-use change and management. For Annex-I countries under the UNFCCC, these data are drawn from the annual GHG inventories submitted to the UNFCCC by each country; for non-Annex-I countries, data are drawn from the most recently submitted National Communication where available. Because of differences in reporting years and methodologies, these data are not generally considered comparable across countries. Data are in million metric tons.

18. Government expenditure on education, total (% of government expenditure) - percentage of total general government expenditure on all sectors (including health, education, social services, etc.). It includes expenditure funded by transfers from international sources to government. General government usually refers to local, regional and central governments.

19. Heat Index 35 (projected change in days) - Total count of days per year where the daily mean Heat Index rose above 35°C. The Heat Index is a measure of how hot it really feels when relative humidity is factored in with the actual air temperature.

20. Individuals using the Internet (% of population) - individuals who have used the Internet (from any location) in the last 3 months. The Internet can be used via a computer, mobile phone, personal digital assistant, games machine, digital TV etc.

21. Income share held by lowest 20% - the share that accrues to subgroups of population indicated by deciles or quintiles. Percentage shares by quintile may not sum to 100 because of rounding.

22. Life expectancy at birth, total (years) - the number of years a newborn infant would live if prevailing patterns of mortality at the time of its birth were to stay the same throughout its life.

23. Literacy rate, adult total (% of people ages 15 and above) - percentage of people ages 15 and above who can both read and write with understanding a short simple statement about their everyday life.

24. Mean Drought Index (projected change, unitless) - The Standardized Precipitation Evapotranspiration Index (SPEI), or Mean Drought Index, calculated for a 12-month period, has been found to be closely related to drought impacts on ecosystems, crop, and water resources. The SPEI is designed to take into account both precipitation and potential evapotranspiration (PET) in determining drought.

25. Methane emissions (metric tons of CO2 equivalent per capita) - those stemming from human activities such as agriculture and from industrial methane production.

26. Mortality rate, under-5 (per 1,000 live births) - probability per 1,000 that a newborn baby will die before reaching age five, if subject to age-specific mortality rates of the specified year.

27. Net migration - Net migration is the net total of migrants during the period, that is, the total number of immigrants less the annual number of emigrants, including both citizens and noncitizens. Data are five-year estimates.

28. Nitrous oxide emissions (metric tons of CO2 equivalent per capita) - emissions from agricultural biomass burning, industrial activities, and livestock management.

29. People using safely managed drinking water services (% of population) - The percentage of people using drinking water from an improved source that is accessible on premises, available when needed and free from faecal and priority chemical contamination. Improved water sources include piped water, boreholes or tubewells, protected dug wells, protected springs, and packaged or delivered water.

30. People using safely managed sanitation services (% of population) - The percentage of people using improved sanitation facilities that are not shared with other households and where excreta are safely disposed of in situ or transported and treated offsite. Improved sanitation facilities include flush/pour flush to piped sewer systems, septic tanks or pit latrines: ventilated improved pit latrines, compositing toilets or pit latrines with slabs.

31. PM2.5 air pollution, mean annual exposure (micrograms per cubic meter) - the average level of exposure of a nation's population to concentrations of suspended particles measuring less than 2.5 microns in aerodynamic diameter, which are capable of penetrating deep into the respiratory tract and causing severe health damage. Exposure is calculated by weighting mean annual concentrations of PM2.5 by population in both urban and rural areas.

32. Population density (people per sq. km of land area) - midyear population divided by land area in square kilometers. Population is based on the de facto definition of population, which counts all residents regardless of legal status or citizenship--except for refugees not permanently settled in the country of asylum, who are generally considered part of the population of their country of origin. Land area is a country's total area, excluding area under inland water bodies, national claims to continental shelf, and exclusive economic zones. In most cases the definition of inland water bodies includes major rivers and lakes.

33. Prevalence of overweight (% of adults) - percentage of adults ages 18 and over whose Body Mass Index (BMI) is more than 25 kg/m2. Body Mass Index (BMI) is a simple index of weight-for-height, or the weight in kilograms divided by the square of the height in meters.

34. Prevalence of undernourishment (% of population) - percentage of the population whose food intake is insufficient to meet dietary energy requirements continuously. Data showing as 5 may signify a prevalence of undernourishment below 5%.
Proportion of seats held by women in national parliaments (%) - the percentage of parliamentary seats in a single or lower chamber held by women.

35. Ratio of female to male labor force participation rate (%) (modeled ILO estimate) - Labor force participation rate is the proportion of the population ages 15 and older that is economically active: all people who supply labor for the production of goods and services during a specified period. Ratio of female to male labor force participation rate is calculated by dividing female labor force participation rate by male labor force participation rate and multiplying by 100.

36. Renewable energy consumption (% of total final energy consumption) - share of renewable energy in total final energy consumption.
Renewable electricity output (% of total electricity output) - the share of electricity generated by renewable power plants in total electricity generated by all types of plants.

37. Research and development expenditure (% of GDP) - Gross domestic expenditures on research and development (R&D), expressed as a percent of GDP. They include both capital and current expenditures in the four main sectors: Business enterprise, Government, Higher education and Private non-profit. R&D covers basic research, applied research, and experimental development.

38. School enrollment, primary (% gross) - Gross enrollment ratio is the ratio of total enrollment, regardless of age, to the population of the age group that officially corresponds to the level of education shown. Primary education provides children with basic reading, writing, and mathematics skills along with an elementary understanding of such subjects as history, geography, natural science, social science, art, and music.

39. Unmet need for contraception (% of married women ages 15-49) - the percentage of fertile, married women of reproductive age who do not want to become pregnant and are not using contraception.

40. Unemployment, total (% of total labor force) (modeled ILO estimate) - the share of the labor force that is without work but available for and seeking employment.






### Step 2: Tidy up the dataset

```{r}
library(tidyverse)
library(dplyr)

#tidy up the first data set

se_clean <- se %>%
  # get rid of unnecessary or redundant variable
  select (-country_code,-Time_code) %>%
  # rename some variables 
  rename("year" = Time, "country" = ï..Country_Name  ) %>%
  # change all the non-numerical variable to NA
  mutate_all(na_if, "..")

#check se data
glimpse (se_clean)
dim(se_clean)



#tidy up the second data set

esg_clean <- esg %>%
  #get rid of unnecessary or redundant variable
  select(-Country_Code, -Series_Code) %>%
  #tidy up the year variable
  pivot_longer(cols=contains('YR'), names_to = "year", values_to = "value") %>%
  separate(year, into=c("year_num", "year_code"), sep="X|\\..|\\.") %>%
  select(-year_num)%>%
  #rename some variable to common variable
  rename("year" = year_code, "country"= ï..Country_Name)

# tidy up the rest of the indicator variables
esg_clean <- esg_clean %>%
  # separate the column series_name to get rid of unit
  separate(Series_Name, into=c("indicator","Unit"), sep="\\(|\\)") %>%
  select(-Unit)%>%
  # pivot wider so that each indicator is a unique variable
  pivot_wider(names_from=indicator, values_from= value) %>%
  #to match up the year between 1990 and 2016
  filter(between(year, 1990, 2016))%>%
  # delete any variables with too many NA or no data
  select(-'Cooling_Degree_Days ', -'Heat_Index_35 ', -'Mean_Drought_Index ', -'Unmet_need_for_contraception ') %>%
  # have all non-numerical values to NA other than country variable
  mutate_all(na_if, "..")
  
# combine variables that are technically same
esg_clean <- esg_clean %>%
  unite('Annual_freshwater_withdrawals_total', 'Annual_freshwater_withdrawals_total ',
        'Annual_freshwater_withdrawals _total ', sep = "", na.rm=T) %>%
  unite('Agriculture_forestry_and_fisihing_value_added', 'Agriculture_forestry_and_fishing_value_added ',
        'Agriculture _forestry _and_fishing _value_added ', sep="", na.rm=T)%>%
  unite('Droughts_floods_extreme_temperatures', 'Droughts_floods_extreme_temperatures ',
        'Droughts _floods _extreme_temperatures ', sep="", na.rm=T) %>%
  unite('Energy_imports_net', 'Energy_imports_net ', 'Energy_imports _net ', sep="", na.rm=T) %>%
  unite('Fertility_rate_tot', 'Fertility_rate_total ', 'Fertility_rate _total ', sep="", na.rm=T) %>%
  unite('Research_development_expenditure', 'Research_and_development_expenditure ',
        'research_and_development_expenditure ', sep="", na.rm=T)

esg_clean <- esg_clean %>%
  mutate_all(na_if, "")

#to check %NA of each variable 
map(esg_clean, ~mean(is.na(.)))


#get rid of variables with more than 30% NA or any insignificant variables
esg_clean <-  esg_clean %>%
  select(-Annual_freshwater_withdrawals_total, -Droughts_floods_extreme_temperatures,
         -`Literacy_rate _adult_total `, -Net_migration, -`Prevalence_of_undernourishment `,
         - `GHG_net_emissions/removals_by_LUCF `,-`Government_expenditure_on_education _total `,
         - `Income_share_held_by_lowest_20%`, -`People_using_safely_managed_drinking_water_services `,
         - `People_using_safely_managed_sanitation_services `,-`PM2.5_air_pollution _mean_annual_exposure `)

#check esg data
glimpse (esg_clean)
dim(esg_clean)
  
```
*After tidying the data, the first data set 'se' has 270 rows and 11 columns while the second data set 'esg' has 270 rows and 28 columns.*





### Step 3: Join or Merge two datasets as one

```{r}
#combine both data by country and year
sesg <- merge(se_clean, esg_clean, by=c("country", "year"))

sesg <- sesg %>%
  # change numerical variable to numeric from character
  mutate_at (vars(-country),as.numeric) %>%
  #rename to shorten for convenience
    rename("acc_cft_t" = access_cft_tot,"acc_elct_t"= access_electricity_tot, "re_elect_out"=ren_electricity_output,
         "re_consum"=re_consumption, "elect_out_t"= total_electricity_output, "re_elect_prcnt"=share_re_in_electricity,
         "e_consum_t"=total_final_energy_consum, "re_consum_prcnt" = share_total_re_in_TEFC,
         "e_int_prim" = primary_energy_intensity, "nat_res_depl"=`natural_resources_depletion `,
         "net_for_depl"=`net_forest_depletion `, "agri_land"=`Agricultural_land `, 
         "agri_for_fish_val" = Agriculture_forestry_and_fisihing_value_added, "ctrl_corrupt"=Control_of_Corruption_Estimate,
         "elect_prd_coal_prcnt" =`Electricity_production_from_coal_sources `, "e_import_net" = Energy_imports_net,
         "e_use"=`Energy_use `, "fertil_r_t" = Fertility_rate_tot, "food_prd" = `Food_production_index `, 
         "forest_a" = `Forest_area `, "GDP_r"=`GDP_growth `, "fssil_e_consum_prcnt" = `Fossil_fuel_energy_consumption `,
         "Intrnt_prcnt"=`Individuals_using_the_Internet `, "life_expct" = `Life_expectancy_at_birth _total `,
         "mortality_r_5-" = `Mortality_rate _under-5 `, "pop_dens"= `Population_density `, 
         "prev_ovrweight" = `Prevalence_of_overweight `, "rd_expend" = Research_development_expenditure,
         "women_legis_prcnt"=`Proportion_of_seats_held_by_women_in_national_parliaments `,  
         "ftom_laborF_r"= `Ratio_of_female_to_male_labor_force_participation_rate `,
         "prim_schl_enrll"= `School_enrollment _primary `, "unemply_t" = `Unemployment _total `)

#check the new combined data
head(sesg)
dim(sesg)
```

*After combining two datasets, there are 270 rows and 37 columns in total. *





### Step 4: Create summary statistics

```{r}
# 6 core dplyr functions (filter, select, arrange, group_by, mutate, summarize)

# new numerical variable summary
sesg %>% 
  filter (country != "World", year==2014) %>%
  mutate(ren_consumToout = re_consum*0.2777778/re_elect_out, na.rm=T) %>%
  select(country, ren_consumToout, GDP_r) %>%
  arrange(desc(GDP_r))%>%
  summarize(min_renR = min(ren_consumToout), mean_renR = mean(ren_consumToout),
            median_renR = median(ren_consumToout), max_renR = max(ren_consumToout),
            var_renR = var(ren_consumToout), sd_renR = sd(ren_consumToout))



#categorical variable summary + numerical variable
sesg%>%
  filter(country != "World") %>%
  mutate(GDP_share = case_when(country %in% c("United States","Japan","Italy") ~"high",
                               country %in% c("Korea, Rep.","Australia","Spain") ~ "mid",
                               country %in% c("Greece","Thailand", "Singapore") ~ "low")) %>%
  group_by(GDP_share)%>%
  summarize(avg_diff_bt_fuel_re_consum_prcnt = mean(fssil_e_consum_prcnt-re_consum_prcnt,na.rm=T),
            avg_fssil_e_consum = mean(fssil_e_consum_prcnt, na.rm=T),
            avg_re_consum_prcnt = mean(re_consum_prcnt, na.rm=T)) %>%
  arrange(desc(avg_diff_bt_fuel_re_consum_prcnt))



#Overall summary of All numerical variables
sesg_num <- sesg %>%
  select_if(is.numeric) %>%
  # deselect non-numerically significant variables
  select(-year,-`Methane_emissions `,-`CO2_emissions `, -`Nitrous_oxide_emissions `, -re_consum)

sesg_sd <- sesg_num %>%
  summarize_all(sd,na.rm=T)%>%
  pivot_longer(cols=c(acc_cft_t:unemply_t), names_to="Variables", values_to="sd")
sesg_sd <-data.frame(sesg_sd)

library(stringr)
remove_all_ws<- function(string){
    return(gsub(" ", "", str_squish(string)))
}

summary_table <-data.frame(summary(sesg_num))
summary_table <- summary_table %>%
  select(-Var1) %>%
  rename("Variables" = Var2)%>%
  mutate_if(is.character, remove_all_ws)%>%
  separate(Freq, into=c("indicator","stats"), sep=":") %>%
  pivot_wider(names_from = indicator, values_from = stats)%>%
  select(-"NA") %>%
  mutate_at("NA's",replace_na, 0)

stat_summary <- merge(summary_table, sesg_sd, by="Variables")
library(kableExtra)
stat_summary %>%
  kbl() %>%
  kable_styling()



# correlation across numerical variables
sesg_cor <- data.frame(cor(sesg_num, use = "pairwise.complete.obs"))

sesg_cor %>%
  mutate_if(is.numeric, ~cell_spec(.x,"html", color = ifelse(.x < 0, "red"," black")))%>%
  kable(format = "html", escape = FALSE) %>%
  kable_styling("striped", full_width = FALSE)

```





### Step 5: Make Visualizations
```{r}
# correlation heatmap - see the last page to see better heat map with values
cor(sesg_num, use = "pairwise.complete.obs") %>%
  # Save as a data frame
  as.data.frame %>%
  # Convert row names to an explicit variable
  rownames_to_column %>%
  # Pivot so that all correlations appear in the same column
  pivot_longer(-1, names_to = "other_var", values_to = "correlation") %>%
  ggplot(aes(rowname, other_var, fill=correlation)) +
  # Heatmap with geom_tile
  geom_tile() +
    theme(axis.text.x=element_text(angle=90,hjust=1))+
  # Change the scale to make the middle appear neutral
  scale_fill_gradient2(low="red",mid="white",high="blue") +
  # Overlay values
  geom_text(aes(label = round(correlation,2)), color = "black", size = 3) +
  # Give title and labels
  labs(title = "Correlation matrix for the dataset sesg", x = "variable 1", y = "variable 2")



# plot 1
sesg_clean <- sesg %>%
  filter(country !="World") %>%
  select(-`Methane_emissions `, -`CO2_emissions `, -`Nitrous_oxide_emissions `)%>%
  mutate(GDP_share = case_when(country %in% c("United States","Japan","Italy") ~"high",
                               country %in% c("Korea, Rep.","Australia","Spain") ~ "mid",
                               country %in% c("Greece","Thailand", "Singapore") ~ "low"))

ggplot(sesg_clean, aes(x=re_consum_prcnt, y=ctrl_corrupt, color= country, shape= GDP_share))+
  geom_jitter()+
  geom_violin(draw_quantiles=0:4/4, trim=FALSE)+
  labs(title="Renewable energy consumption percentage vs. Govt. corruption estimate",
       x= "Renewable Energy Consumption percentage (%)",
       y="Government Corruption Estimate")
        
 

# plot 2
ggplot(sesg_clean, aes(x=re_consum_prcnt, y= agri_for_fish_val, color=GDP_share))+
  geom_point()+
  geom_boxplot()+
  labs(title="Renewable Energy Consumption percentage vs. Agriculture forest and fishing value added", 
       x= "Renewable Energy Consumption percentage (%)",
       y= "Agriculture, value added (% of GDP)")

```

*Geom_point is added to the graph to show the actual data points and ranges that box plots cannot show.*

*Variables "renewable energy consumption percentage" and "control of corruption" are selected because they show negative correlation to one another (-0.64).  The first ggplot shows that government corruption estimate decreases as renewable energy consumption percentage increases. Colors represent country and the shape of each data points shows the GDP share that each country belongs. GDP share follows ordinal manner with these variables. *

*Variables "agriculture (fishing and forest), value added, and renewable energy consumption percentage are selected because they show positive correlation to one another (0.77). The second ggplot shows that as renewable energy consumption percentage increases the agriculture, value added percentage also increases. Colors represent GDP share of each country. It is clear that GDP does not show any significant relationship with any variables. Low GDP share countries has the greatest variance.*





### Step 6: Perform k-means/PAM clustering or PCA
```{r}
#clean the data
sesg_clean <- sesg_clean %>%
  na.omit
  
#scale
sesg_scaled <- sesg_clean %>%
  select(where(is.numeric))%>%
  scale

#peform PCA
sesg_pca <- sesg_scaled %>%
  prcomp()
sesg_pca
summary(sesg_pca)

#screeplot
library(factoextra)
fviz_screeplot(sesg_pca, addlabels=T)

sesg_pca_data <- data.frame (sesg_pca$x, country=sesg_clean$country, year= sesg_clean$year, GDP_share=sesg_clean$GDP_share)
sesg_pca_data
ggplot(sesg_pca_data, aes(x = PC1, y = PC2, color = country, shape=GDP_share)) + 
  geom_point() +
  labs(title="PC1 vs. PC2")

#each variable contribution to each PC
sesg_pca$rotation
fviz_pca_var(sesg_pca, col.var = "black")

# the ideal number of clustering
library(cluster)
sesg_pca_data %>%
  select(where(is.numeric)) %>%
  fviz_nbclust(FUNcluster = pam, method = "s")

# pam algorithm
sesg_pam <- sesg_pca_data %>%
  select(PC1,PC2) %>%
  pam(k=3)
sesg_pam

#clustering
sesg_pamclust <- sesg_pca_data %>%
  mutate(cluster=as.factor(sesg_pam$clustering))
sesg_pamclust

# graph with clusters
sesg_pamclust %>% 
  ggplot(mapping = aes(x= PC1, y= PC2)) +
  geom_point(aes(color=country, shape= GDP_share))+
  stat_ellipse(aes(group=cluster))
```
*By the rule of cumulative proportion being 80%, PC1 through PC5 are considered, while PC1 through PC7 need to be considered by Kaiser's rule where eigenvalues are greater than 1.*

*primary energy intensity, female to male ratio in labor force, fertility rate, renewable energy consumption, renewable energy output, control of corruption, internet user percentage, research and development expenditure, electricity production from fossil fuel, energy use, fossil fuel energy consumption percentage are positively contributed to both PC1 and PC2 *

*k=3 is selected because the GDP share is divided into 3 categories. It wouldn't make sense to choose k=7*





### Main Findings
*The first main finding: Some environmental-related variables show some correlation to social-related variables. For example, life expectancy has a positive correlation with access to clean fuel and technology in total (0.80). Net forest depletion has a negative correlation with access to clean fuel and technology (-0.93). *

*The second main fining: It shows that GDP rate does not have any strong correlation with amount of renewable energy (-0.12). In addition GDP share does not show any significant correlation with other sustainability-related factors.*


```{r}
Sys.info()
```
