---
title: "Project 2: Airquality seasonality"
author: "Hye Jeong Lee"
date: "4/13/2021"
always_allow_html: yes
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(plotROC)
```

### 1. Introduction 
```{r}
# dataset
aq<- airquality

# dimension of dataset
dim(aq)

# change numerical month to categorical month
aq$month <- month.abb[aq$Month]
aq <- aq %>%
  select(-Month) %>%
  rename ("Month" = month) %>%
  mutate(Temp=(aq$Temp -32)*(5/9))

```
The data set 'air quality' is one of built-in R studio which contains 153 observations and 5 variables. Month is the only categorical variable and the rests are numeric variables.

The purpose of this analysis is to test if there is a significant difference in means of ozone concentration between months, and to build a linear regression model to predict ozone based on variables as well as to build a logistic regression model that predict if the ozone concentration is permissible based on other variables.  

I expect to find that there is a significant difference in mean of ozone concentration between months, and there is a significant effect of solar radiation and month on ozone concentration. In addition, I expect to find that there is a significant effect of solar radiation, month, temperature on predicting if the ozone concentration is permissible ( < 100 ppb).  

I chose this dataset because everyone on earth is affected by the outside air quality and air quality related diseases are increasing with time. Finding any pattern or seasonality in the air quality would be important for improving air quality. 

__Variables__

1. ozone: mean ozone in parts per billion from 1300 to 1500 hours at Roosevelt Island (ppb)

2. Solar.R: solar radiation in Langleys in frequency band 4000 - 7700 Angstroms from 0800 to 1200 hours at central park (lang)

3. wind: average wind speed in miles per hour at 0700 and 1000 hours at LaGuardia Airport (mph)

4. temp: maximum daily temperature in degrees Farenheit at La Guardia Airport (Farenheit)

5. month: abbreviated character Month

6. day: numeric day of month (1-31)



### 2. EDA
```{r}
# correlation heatmap to visualize overall correlation to pick variables
cor(airquality, use = "pairwise.complete.obs") %>%
  # Save as a data frame
  as.data.frame %>%
  # Convert row names to an explicit variable
  rownames_to_column %>%
  # Pivot so that all correlations appear in the same column
  pivot_longer(-1, names_to = "other_var", values_to = "correlation") %>%
  ggplot(aes(rowname, other_var, fill=correlation)) +
  # Heatmap with geom_tile
  geom_tile() +
    theme(axis.text.x=element_text(angle=90,hjust=1))+
  # Change the scale to make the middle appear neutral
  scale_fill_gradient2(low="red",mid="white",high="blue") +
  # Overlay values
  geom_text(aes(label = round(correlation,2)), color = "black", size = 5) +
  # Give title and labels
  labs(title = "Correlation matrix for the dataset airquality", x = "variable 1", y = "variable 2")


# summary of each variables  into a table
aq_sd <- airquality %>%
  summarize_all(sd,na.rm=T)%>%
  pivot_longer(cols=c(Ozone:Day), names_to="Variables", values_to="sd")
aq_sd <-data.frame(aq_sd)
## remove white space
library(stringr)
remove_all_ws<- function(string){
    return(gsub(" ", "", str_squish(string)))
}
## summary 
summary_tbl <-data.frame(summary(airquality))
summary_tbl <- summary_tbl %>%
  select(-Var1) %>%
  rename("Variables" = Var2)%>%
  mutate_if(is.character, remove_all_ws)%>%
  separate(Freq, into=c("indicator","stats"), sep=":") %>%
  pivot_wider(names_from = indicator, values_from = stats)%>%
  select(-"NA")%>%
  mutate_at("NA's",replace_na, 0)
## combine
stat_summary <- merge(summary_tbl, aq_sd, by="Variables")
library(kableExtra)
## pretty table
stat_summary %>%
  kbl() %>%
  kable_styling()


# relationship graph for MANOVA
library(psych)
pairs.panels(airquality, 
             method = "pearson", # correlation coefficient method
             hist.col = "blue", # color of histogram 
             smooth = FALSE, density = FALSE, ellipses = FALSE)

m <- c("May","June","July","August","September")


# specific univariate and bivariate graphs
aq1 <- aq %>% mutate(Month = factor(Month, 
                    levels = c("May", "Jun", "Jul", "Aug", "Sep"),
                    ordered = TRUE))
## ozone by month
ggplot(aq1, aes(x=Month,y=Ozone, fill=Month))+
  geom_boxplot()+
  labs(title="Ozone concentration by Month", y="Ozone (ppb)")
## relationship between ozone and solar radiation
ggplot(aq, aes(x=Ozone, y=Solar.R))+
  geom_point()+
  geom_smooth(method=lm,se=F)+
  labs(title="Ozone vs. Solar Radiation", x="Ozone concentration (ppb)", y="Solar Radiation (langley)")
## relationship between ozone and temperature
ggplot(aq, aes(x=Ozone, y=Temp))+
  geom_point()+
  geom_smooth(method=lm,se=F)+
  labs(title="Ozone vs. Temperature", x= "Ozone concentration (ppb)", y="Temperature (Celsius)")
## relationship between ozone and Wind speed
ggplot(aq, aes(x=Ozone, y=Wind))+
  geom_point()+
  geom_smooth(method=lm,se=F)+
  labs(title="Ozone vs. Wind Speed", x= "Ozone concentration (ppb)", y="Wind Speed (mph)")
```

The correlation heatmap shows that there is a strong positive relationship between temperature and ozone concentration. There is a strong negative correlation between wind speed and ozone concentration. Ozone and speed have some negative relationship.

Statistical analysis are done for each variable including minimum, maximum, interquartile, average and standard deviation along with number of NAs in the data.

Based on the bar graph, July and August have larger variance and range in ozone concentration in ppb compared to the other months. 
Based on the line graph, ozone concentration in ppb and solar radiation in langley have positive relationship to one another. It would be more fitting if the line was curved as square root function.
Ozone concentration in ppb and temperature in Celsius have positive relationship to one another in a linear fashion. 
Ozone concentration in ppb and wind speed in mph have negative relationship to one another in a linear manner. 


### 3. MANOVA
```{r}
#Check for assumption
## 1st assumption - normality
aq1 %>%
  group_by(Month) %>%
  summarize(p.value=shapiro.test(Ozone)$p.value)
## data transformation 
aq1 <- aq1 %>%
  mutate (lnozone = log(Ozone))
ggplot(aq1, aes(x=Month,y=lnozone, fill=Month))+
  geom_boxplot()+
  labs(title="Ozone concentration by Month", y="log Ozone (logppb)")
aq1 %>%
  group_by(Month) %>%
  summarize(p.value=shapiro.test(lnozone)$p.value)
## 2nd assumption - equal variance
aq1 %>%
  group_by(Month) %>%
  summarize(variance = var(lnozone,na.rm=T))

# anova
summary(aov(lnozone ~ Month, data = aq1))
pairwise.t.test(aq1$lnozone, aq1$Month, p.adj = "none")
pairwise.t.test(aq1$lnozone, aq1$Month, p.adj = "bonferroni")

# manova
manova_aq <- manova(cbind(lnozone,Temp,Solar.R, Wind) ~ Month, data = aq1)
summary(manova_aq) ## significant p < 0.001
 ## one-way ANOVA for each
 summary.aov(manova_aq) # lnozone, temp, wind are significant
  ### post-hoc analysis - for lnozone
  pairwise.t.test(aq1$lnozone, aq1$Month, p.adj="none")
  ### post-hoc analysis - for temp
  pairwise.t.test(aq1$Temp, aq1$Month, p.adj="none")
  ### post-hoc analysis - for wind
  pairwise.t.test(aq1$Wind, aq1$Month, p.adj="none")

  ### post-hoc analysis - for lnozone with correction
  pairwise.t.test(aq1$lnozone, aq1$Month, p.adj="bonferroni")
  ### post-hoc analysis - for temp with correction
  pairwise.t.test(aq1$Temp, aq1$Month, p.adj="bonferroni")
  ### post-hoc analysis - for wind with correction
  pairwise.t.test(aq1$Wind, aq1$Month, p.adj="bonferroni")
  
# permanova
  library(vegan)
 ## Compute Euclidean distances between observations
 dists <- aq1 %>%
  select(lnozone, Temp, Wind) %>%
  dist
 ## Perform PERMANOVA on the distance matrix
 adonis(dists ~ Month, data = aq1)
```
After the data transformation on ozone concentration, the normality and homodescedasity assumptions are met.

The null hypothesis is that there is no significant difference in means of natural log of ozone concentration between months. 
The alternative hypothesis is that there is significant difference in means of natural log of ozone concentration between at least one month and the rest.
*p-value of anova test is small enough to reject the null (p<0.001). There is a significant difference in ln ozone between months. With Bonferroni correction, there is a significant difference in ln ozone between May and July, May and August, July and September, and August and September.*

The null hypothesis is that there is no significant difference in means of each variable between months. The alternative hypothesis is that there is significant difference in means of each variable between at least one month and the rest.
*p-value of manova test is small enough to reject the null. There is a significant difference in each variable for each month. All months are significantly different in natural log of ozone concentration (p<0.001), temperature (p<0.001) and wind (p=0.004).*

Bonferonni alpha = 0.05/(5+5*3)= 0.05/20 = 0.0025

With Bonferonni correction, May and July, May and August, July and September, and August and September pairs in natural log of ozone concentration appear to be significantly different at alpha = 0.05.
With Bonferonni correction, May and June, May and July, May and August, May and September, June and August, July and September, August and September pairs in temperature appear to be significantly different at alpha = 0.05.
With Bonferonni correction, May and July, May and August pairs in wind speed appear to be significantly different at alpha = 0.05

From permanova, it is clear that lnozone, temperature and wind speed are differ across at least Month.



### 4. Randomization Test
```{r}
set.seed(348)
aq4 <- aq1 %>%
  filter(!is.na(lnozone))

# LNOZONE
summary(aov(lnozone ~ Month, data = aq4))
pairwise.t.test(aq1$lnozone, aq1$Month, p.adj = "none")
obs_F <- 9.163
## randomization test
Fs <- replicate(1000,{
  # Randomly permute the response variable across doses
  new <- aq4 %>%
    mutate(lnozone = sample(lnozone))
  # Compute variation within groups
  SSW <- new %>%
    group_by(Month) %>%
    summarize(SSW = sum((lnozone - mean(lnozone,na.rm=T))^2)) %>%
    summarize(sum(SSW)) %>% 
    pull
  # Compute variation between groups
  SSB <- new %>% 
    mutate(mean = mean(lnozone, na.rm=T)) %>%
    group_by(Month) %>% 
    mutate(groupmean = mean(lnozone,na.rm=T)) %>%
    summarize(SSB = sum((mean - groupmean)^2)) %>%
    summarize(sum(SSB)) %>%
    pull
  # Compute the F-statistic (ratio of MSB and MSW)
  # df for SSB is 5 groups - 1 = 4
  # df for SSW is 153 observations - 5 groups = 148
  (SSB/4)/(SSW/148)
  })
 ## Represent the distribution of the F-statistics for each randomized sample
 hist(Fs, prob=T); abline(v = obs_F, col="red",add=TRUE)
 ## Calculate the proportion of F statistic that are greater than the observed F-statistic - reject the null hypothesis
 mean(Fs > obs_F)

 # TEMPERATURE
summary(aov(Temp ~ Month, data = aq1))
pairwise.t.test(aq1$Temp, aq1$Month, p.adj = "none")
obs_F <- 39.85
## randomization test
Fs <- replicate(1000,{
  # Randomly permute the response variable across doses
  new <- aq1 %>%
    mutate(Temp = sample(Temp))
  # Compute variation within groups
  SSW <- new %>%
    group_by(Month) %>%
    summarize(SSW = sum((Temp - mean(Temp,na.rm=T))^2)) %>%
    summarize(sum(SSW)) %>% 
    pull
  # Compute variation between groups
  SSB <- new %>% 
    mutate(mean = mean(Temp, na.rm=T)) %>%
    group_by(Month) %>% 
    mutate(groupmean = mean(Temp,na.rm=T)) %>%
    summarize(SSB = sum((mean - groupmean)^2)) %>%
    summarize(sum(SSB)) %>%
    pull
  # Compute the F-statistic (ratio of MSB and MSW)
  # df for SSB is 5 groups - 1 = 4
  # df for SSW is 153 observations - 5 groups = 148
  (SSB/4)/(SSW/148)
  })
 ## Represent the distribution of the F-statistics for each randomized sample
 hist(Fs, prob=T); abline(v = obs_F, col="red",add=T)
 ## Calculate the proportion of F statistic that are greater than the observed F-statistic - reject the null hypothesis
 mean(Fs > obs_F)
 
 # WIND
summary(aov(Wind ~ Month, data = aq1))
pairwise.t.test(aq1$Wind, aq1$Month, p.adj = "none")
obs_F <- 3.529
## randomization test
Fs <- replicate(1000,{
  # Randomly permute the response variable across doses
  new <- aq1 %>%
    mutate(Wind = sample(Wind))
  # Compute variation within groups
  SSW <- new %>%
    group_by(Month) %>%
    summarize(SSW = sum((Wind - mean(Wind,na.rm=T))^2)) %>%
    summarize(sum(SSW)) %>% 
    pull
  # Compute variation between groups
  SSB <- new %>% 
    mutate(mean = mean(Wind, na.rm=T)) %>%
    group_by(Month) %>% 
    mutate(groupmean = mean(Wind,na.rm=T)) %>%
    summarize(SSB = sum((mean - groupmean)^2)) %>%
    summarize(sum(SSB)) %>%
    pull
  # Compute the F-statistic (ratio of MSB and MSW)
  # df for SSB is 5 groups - 1 = 4
  # df for SSW is 153 observations - 5 groups = 148
  (SSB/4)/(SSW/148)
  })
 ## Represent the distribution of the F-statistics for each randomized sample
 hist(Fs, prob=T); abline(v = obs_F, col="red",add=T)
 ## Calculate the proportion of F statistic that are greater than the observed F-statistic - reject the null hypothesis
 mean(Fs > obs_F)


```  

1. The null hypothesis is that the mean of natural log of ozone concentration is same across Month. The alternative hypothesis is that mean of natural log of ozone concentration is different across Month at least one. There is a significant difference in mean of natural log of ozone concentration across Month (F=9.163, df=4, p < 0.001). This concurs with the result from the randomization test.(p=0.001)

2. The null hypothesis is that the mean of temperature is same across Month. The alternative hypothesis is that mean of temperature is different across Month. There is a significant difference in mean of temperature across Month at least one (F=39.85, df=4, p < 0.001). This concurs with the result from the randomization test.(p=0)

3. The null hypothesis is that the mean of wind speed is same across Month. The alternative hypothesis is that mean of wind speed is different across months at least one. There is a significant difference in mean of wind speed across months (F=3.529, df=4, p < 0.01). This does not concur but still holds the significance with the result from the randomization test.(p=0.013)



### 5. Linear Regression model
```{r}
library(lmtest)
library(sandwich)
set.seed(348)

aq2 <- aq1 %>%
  mutate (lnozone_c = lnozone - mean(lnozone, na.rm=T),
          Temp_c = Temp - mean(Temp, na.rm=T),
          Wind_c = Wind - mean(Wind, na.rm=T))%>%
  mutate (Month = ntile(Month, 5)) %>%
  mutate (Month = factor(Month, labels=c("May","Jun","Jul","Aug", "Sep")))

# linear regression model to predict lnozone based on temperature only
fit <- lm(lnozone_c ~ Temp_c, data = aq2)
summary(fit)
 ## graph
 ggplot(aq2, aes(x=Temp_c, y=lnozone_c))+ geom_point() + geom_smooth(method=lm, se=F) +
   labs(title="scaled temperature vs. scaled ln(ozone)", x="Temperature(C)", y="ln(ozone) (lnppb)")
 ## proportion of variance in lnozone explained by temp_c
 summary(fit)$r.sq
 ## 1st assumption - normality
 shapiro.test(fit$residuals)
 ## 2nd assumption - homoscedasticity
 plot(fit, which = 1)
 bptest(fit)
 ## robust SEs
 coeftest(fit, vcov = vcovHC(fit))
 ## bootstrap SEs
 samp_SEs <- replicate(1000, {
  # Bootstrap your data (resample observations)
  boot_data <- sample_frac(aq2, replace = TRUE)
  # Fit regression model
  fitboot <- lm(lnozone_c ~ Temp_c, data = boot_data)
  # Save the coefficients
  coef(fitboot)
  })
 samp_SEs %>%
  # Transpose the obtained matrices
  t %>%
  # Consider the matrix as a data frame
  as.data.frame %>%
  # Compute the standard error (standard deviation of the sampling distribution)
  summarize_all(sd) #int 0.3098 temp 0.011
 
# linear regression model to predict lnozone based on wind speed only
fit<- lm(lnozone_c ~ Wind_c, data = aq2)
summary(fit)
 ## graph
 ggplot(aq2, aes(x=Wind_c, y=lnozone_c))+ geom_point() + geom_smooth(method=lm, se=F) + 
   labs(title="scaled wind speed vs. scaled ln(ozone)", x="wind speed(mph)", y="ln(ozone) (lnppb)")
 ## proportion of variance in lnozone explained by wind_c
 summary(fit)$r.sq
 ## 1st assumption - normality
 shapiro.test(fit$residuals)
 plot(fit, which = 2)
 ## 2nd assumption - homoscedasticity
 bptest(fit) 
 plot(fit, which = 1)
 ## robust SEs
 coeftest(fit, vcov = vcovHC(fit))
 ## bootstrap SEs
 samp_SEs <- replicate(1000, {
  # Bootstrap your data (resample observations)
  boot_data <- sample_frac(aq2, replace = TRUE)
  # Fit regression model
  fitboot <- lm(lnozone_c ~ Wind_c, data = boot_data)
  # Save the coefficients
  coef(fitboot)
  })
 samp_SEs %>%
  # Transpose the obtained matrices
  t %>%
  # Consider the matrix as a data frame
  as.data.frame %>%
  # Compute the standard error (standard deviation of the sampling distribution)
  summarize_all(sd)


# linear regression model to predict lnozone based on temperature and Month with interaction
fit <- lm(lnozone_c ~ Temp_c * Month, data=aq2)
summary(fit)
 ## graph
 ggplot(aq2, aes(x=Temp_c, y=lnozone_c, color=Month))+ geom_point() + geom_smooth(method=lm, se=F) +
   labs(title="scaled temperature vs. scaled ln(ozone) by month", x="Temperature(C)", y="ln(ozone) (lnppb)")
 ## proportion of variance in lnozone explained by temp_c and month interaction
 summary(fit)$r.sq
 ## 1st assumption - normality
 shapiro.test(fit$residuals)
 plot(fit, which = 2)
 ## 2nd assumption - homoscedasticity
 bptest(fit) 
 plot(fit, which = 1)
 ## robust SEs
 coeftest(fit, vcov = vcovHC(fit))
 ## bootstrap SEs
 samp_SEs <- replicate(1000, {
  # Bootstrap your data (resample observations)
  boot_data <- sample_frac(aq2, replace = TRUE)
  # Fit regression model
  fitboot <- lm(lnozone_c ~ Temp_c * Month, data = boot_data)
  # Save the coefficients
  coef(fitboot)
  })
 samp_SEs %>%
  # Transpose the obtained matrices
  t %>%
  # Consider the matrix as a data frame
  as.data.frame %>%
  # Compute the standard error (standard deviation of the sampling distribution)
  summarize_all(sd)
 
 # linear regression model to predict lnozone based on wind and Month 
fit <- lm(lnozone_c ~ Wind_c * Month, data=aq2)
summary(fit)
 ## graph
 ggplot(aq2, aes(x=Wind_c, y=lnozone_c, color=Month))+ geom_point() + geom_smooth(method=lm, se=F) +
   labs(title="scaled wind speed vs. scaled ln(ozone)", x="wind speed (mph)", y="ln(ozone) (lnppb)")
 ## proportion of variance in lnozone explained by wind_c and month
 summary(fit)$r.sq
## 1st assumption - normality
 shapiro.test(fit$residuals)
 plot(fit, which = 2)
 ## 2nd assumption - homoscedasticity
 bptest(fit) 
 plot(fit, which = 1)
 ## robust SEs
 coeftest(fit, vcov = vcovHC(fit))
 ## bootstrap SEs
 samp_SEs <- replicate(1000, {
  # Bootstrap your data (resample observations)
  boot_data <- sample_frac(aq2, replace = TRUE)
  # Fit regression model
  fitboot <- lm(lnozone_c ~ Wind_c*Month, data = boot_data)
  # Save the coefficients
  coef(fitboot)
  })
 samp_SEs %>%
  # Transpose the obtained matrices
  t %>%
  # Consider the matrix as a data frame
  as.data.frame %>%
  # Compute the standard error (standard deviation of the sampling distribution)
  summarize_all(sd)
 
```

1. Linear regression model that predicts lnozone from temperature met the normality assumption but not the homoscedasticity assumption. Based on the plot, equal variance does not look too off the straight line. Temperature is significantly associated with lnozone; for every 1 unit increase in temperature, there is 0.122 increase in lnozone (t = 11.7, df = 114, p < 0.001). The proportion of variance in lnozone explained by temperature is 0.547 (54.7%).
*lnozone(hat) = 0.000787 + 0.122(temp_c) with standard error of 0.055 on intercept and 0.012 on temp_c. The robust standard error concurs with bootstrapped standard errors (0.054 on intercept, 0.012 on temp_c).*

2. Linear regression model that predicts lnozone from wind speed did not met the normality assumption but not the homoscedasticity assumption. Wind speed is significantly associated with lnozone; for every 1 unit increase in wind speed, there is 0.130 decrease in lnozone (t = -6.8, df = 114, p < 0.001). The proportion of variance in lnozone explained by wind speed is 0.290 (29.0%).
*lnozone(hat) = -0.012-0.13(Wind_c) with standard error of 0.069 on intercept and 0.019 on Wind_c. The robust standard error concurs with bootstrapped standard errors (0.070 on intercept, 0.019 on wind_c).*

3. Linear regression model that predicts lnozone from temperature and month with interaction did not met the normality assumption but not the homoscedasticity assumption. Temperature is significantly associated with lnozone; for every 1 unit increase in temperature, there is 0.132 increase in lnozone (t = 4.12, df = 106, p < 0.001). Month is not a significant predictor of lnozone. There is no significant interaction between month and temperature. The proportion of variance in lnozone explained by wind speed is 0.583 (58.3%).
*lnozone(hat) = 0.212 + 0.132(Temp_c) with standard error of 0.343 on intercept and 0.054 on Temp_c. The robust standard error is simliar to bootstrapped standard errors (0.33 on intercept, and 0.053 on temp_c).*

4. Linear regression model that predicts lnozone from wind and month with interaction did not met the normality assumption but not the homoscedasticity assumption. Wind speed is no longer significantly associated with lnozone; for every 1 unit increase in wind speed, there is 0.057 decrease in lnozone. Month is significantly associated with lnozone controlling wind. For average wind, ln ozone for June is 0.549 lnppb higher than May (t=2.06, df=106, p<0.001). For average wind, ln ozone for July is 0.732 lnppb higher than May (t=3.57, df=106, p<0.001). For average wind, ln ozone for August is 0.727 lnppb higher than May (t=3.551, df=106, p<0.001). The proportion of variance in lnozone explained by wind speed is 0.442 (44.2%).
*The standard errors are 0.239 on intercept, 0.054 on Wind_c, 0.46 on June, 0.28 on July, 0.27 on August, and 0.26 on September. The robust standard error is similar to bootstrapped standard errors.(0.24 on intercept, 0.056 on wind_c, 0.38 on June, 0.27 on July, 0.28 on August, 0.26 on September)*


### 6. Logistic Regression
```{r}
# create a binary variable
aq3 <- aq1 %>%
  mutate(y = ifelse(Ozone < 100, 1, 0))%>%
  mutate (Month = ntile(Month, 5)) %>%
  mutate (Month = factor(Month, labels=c("May","Jun","Jul","Aug", "Sep"))) %>%
  filter(!is.na(Ozone))
aq3

# regression model to predict if ozone concentration is permissible by Month and wind 
fit <- glm(y ~ Month + Wind, data= aq3, family="binomial") 
summary(fit)
 ## interpret
 exp(coef(fit))
 ## Confusion matrix: compare true to predicted condition
 aq3$prob1 <- predict(fit, type = "response")
 aq3$predicted <- ifelse(aq3$prob1 > .5, "permissible", "not permissible")
 table(true_condition = aq3$y, predicted_condition = aq3$predicted) %>% 
  addmargins
  ## accuracy
  (4+108)/116
  ## sensitivity TPR
  108/109
  ## specificity TNR
  4/7
  ## recall precision PPV
  108/111
 ## density 
 aq3$logit <- predict(fit)
 ggplot(aq3, aes(logit, fill = as.factor(y))) +
  geom_density(alpha = .3) +
  geom_vline(xintercept = 0, lty = 2) +
  labs(title = "Density graph of permissible ozone concetration", fill = "Permissible status, 1 permissible")
 ## Save predicted probabilities
 df <- data.frame(prob=predict(fit, type = "response"),y=aq3$y)
 ## Plot ROC curve
 ROCplot <- ggplot(df) + geom_roc(aes(d = y, m = prob), n.cuts = 0)+labs(title="ROC curve of y~ Month+Wind")
 ROCplot
 ## Calculate AUC
 calc_auc(ROCplot)$AUC

# regression model to predict if ozone concentration is permissible by Month and temperature 
fit <- glm(y ~ Month + Temp, data= aq3, family="binomial") 
summary(fit)
 ## interpret
 exp(coef(fit))
  ## Confusion matrix: compare true to predicted condition
 aq3$prob1 <- predict(fit, type = "response")
 aq3$predicted <- ifelse(aq3$prob1 > .5, "permissible", "not permissible")
 table(true_condition = aq3$y, predicted_condition = aq3$predicted) %>% 
  addmargins
  ## accuracy
  (109)/116
  ## sensitivity TPR
  109/109
  ## specificity TNR
  0/7
  ## recall precision PPV
  0/111
 ## density 
 aq3$logit <- predict(fit)
 ggplot(aq3, aes(logit, fill = as.factor(y))) +
  geom_density(alpha = .3) +
  geom_vline(xintercept = 0, lty = 2) +
  labs(title = "Density graph of permissible ozone concetration",fill = "Permissible status, 1 permissible")
 ## Save predicted probabilities
 df <- data.frame(prob=predict(fit, type = "response"),y=aq3$y)
 ## Plot ROC curve
 ROCplot <- ggplot(df) + geom_roc(aes(d = y, m = prob), n.cuts = 0)+labs(title="ROC curve of y~Month+Temp")
 ROCplot
 ## Calculate AUC
 calc_auc(ROCplot)$AUC

# regression model to predict if ozone concentration is permissible by temperature and wind 
fit <- glm(y ~ Temp + Wind, data= aq3, family="binomial") 
summary(fit)
 ## interpret
 exp(coef(fit))
  ## Confusion matrix: compare true to predicted condition
 aq3$prob1 <- predict(fit, type = "response")
 aq3$predicted <- ifelse(aq3$prob1 > .5, "permissible", "not permissible")
 table(true_condition = aq3$y, predicted_condition = aq3$predicted) %>% 
  addmargins
  ## accuracy
  (2+108)/116
  ## sensitivity TPR
  108/109
  ## specificity TNR
  2/7
  ## recall precision PPV
  108/111
 ## density 
 aq3$logit <- predict(fit)
 ggplot(aq3, aes(logit, fill = as.factor(y))) +
  geom_density(alpha = .3) +
  geom_vline(xintercept = 0, lty = 2) +
  labs(title = "Density graph of permissible ozone concetration",fill = "Permissible status, 1 permissible")
 ## Save predicted probabilities
 df <- data.frame(prob=predict(fit, type = "response"),y=aq3$y)
 ## Plot ROC curve
 ROCplot <- ggplot(df) + geom_roc(aes(d = y, m = prob), n.cuts = 0)+labs(title="ROC curve of y~ Temp+Wind")
 ROCplot
 ## Calculate AUC
 calc_auc(ROCplot)$AUC
 

# regression model to predict if ozone concentration is permissible by Month, temperature and wind 
fit <- glm(y ~ Month + Temp + Wind, data= aq3, family="binomial") 
summary(fit)
 ## interpret
 exp(coef(fit))
  ## Confusion matrix: compare true to predicted condition
 aq3$prob1 <- predict(fit, type = "response")
 aq3$predicted <- ifelse(aq3$prob1 > .5, "permissible", "not permissible")
 table(true_condition = aq3$y, predicted_condition = aq3$predicted) %>% 
  addmargins
  ## accuracy
  (4+108)/116
  ## sensitivity TPR
  108/109
  ## specificity TNR
  4/7
  ## recall precision PPV
  108/111
 ## density 
 aq3$logit <- predict(fit)
 ggplot(aq3, aes(logit, fill = as.factor(y))) +
  geom_density(alpha = .3) +
  geom_vline(xintercept = 0, lty = 2) +
  labs(title = "Density graph of permissible ozone concetration",fill = "Permissible status, 1 permissible")
 ## Save predicted probabilities
 df <- data.frame(prob=predict(fit, type = "response"),y=aq3$y)
 ## Plot ROC curve
 ROCplot <- ggplot(df) + geom_roc(aes(d = y, m = prob), n.cuts = 0)+labs(title="ROC curve of y~ Month+Temp+Wind")
 ROCplot
 ## Calculate AUC
 calc_auc(ROCplot)$AUC
 
```
Build a regression model to predict a binary response. Interpret the coefficients and generate a ROC curve, calculate AUC.

1. Controlling Wind, month is not a significant predictor of ozone permissible status. Wind is a significant predictor. Controlling month, for every 1 unit increase in wind speed, the odds of permissible ozone level changes by a factor of 2.48 (increase by 148%). Accuracy is 96.6%, sensitivity is 99.1%, specificity is 57.1% and recall is 97.3%. AIC = 38.255. AUC = 0.926

2. Controlling temperature, month is not a significant predictor of ozone permissible status. Temperature is a significant predictor. Controlling month, for every 1 unit increase in temperature, the odds of permissible ozone level changes by a factor of 0.757. Accuracy is 94.0%, sensitivity is 100%, specificity is 0% and recall is 0%. AIC = 52.088. AUC = 0.876 

3. Controlling temperature, wind is a significant predictor of ozone permissible status (p<0.01). However, controlling wind, temperature is not a significant predictor. Controlling temperature, for every 1 unit increase in wind the odds of permissible ozone level changes by a factor of 2.18 (increase by 118%). Accuracy is 94.8%, sensitivity is 99.1% , specificty is 28.6% and recall is 89.1%. AIC = 39.72. AUC = 0.891

4. Controlling temperature and month, wind is a significant predictor of ozone permissible status (p < 0.01). Controlling temperature and month, for every 1 unit increase in wind, the odds of permissible ozone level changes by a factor of 2.33 (increase by 133%). Accuracy is 96.6%, Sensitivity = 99.1%, specificity = 0.571, and recall is 97.3%. AIC = 39.453. AUC = 0.941

By AUC, a model with month, wind and temperature does a better job of predicting if the ozone level is permissible or not.
By AIC, a model with month and wind fits better than the other models. 


```{r, echo=F}
## DO NOT DELETE THIS BLOCK!
Sys.info()
```

